\documentclass[serif]{sfuthesis}

%% Fill in below, the values here are fine for a physics masters (just put the name of the thesis and your name and update the date and previous degrees info).

%% No quotes in the title, do NOT end it with a period
\title{Variance in Initiation Factors Does Not Strongly Affect the Replication Profile of Budding Yeast DNA}
\thesistype{Thesis}
\author{Mike Chomitz}
\previousdegrees{%
	B.Sc., Trent University, 2012}
\degree{Master of Science}
\discipline{Physics}
\department{Department of Physics}
\faculty{Faculty of Science}
\copyrightyear{2015}
\semester{Fall 2015}
\date{August 19, 2015}

%% Choose words or phrases that will help people locate your thesis via search tools (library catalogues, Google, etc.)
%% Keywords appear under the abstract and must be entered into the Thesis Registration System.
\keywords{DNA replication kinetics; multiple initiator model; phantom nuclei; Poisson process}


\committee{%
	\chair{Dr.\ Some Body}{POSITION IN DEPARTMENT}
	\member{Dr.\ John Bechhoefer}{Senior Supervisor\\Professor}
	\member{Dr.\ Malcolm Kennet}{Supervisor\\Associate Professor}
	\member{Dr.\ Eldon Emberly}{Internal Examiner\\Associate Professor}
%% add or remove \members as need
}

%  PACKAGES AND CUSTOMIZATIONS  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  You don't need to call the following packages, which are already called
%  in the sfuthesis class file:
%  - etoolbox
%  - setspace
%  - enumitem
%  - pdfpages
%  - tocloft
%  - appendix
%  - cmbright or lmodern
%
%  If you call one of the above packages (or one of their dependencies)
%  with different options, you may get a "Option clash" LaTeX error.
%  If you get this error, you can fix it by removing your copy of
%  \usepackage and just passing the options you need by adding
%
%  \PassOptionsToPackage{<options>}{<package>}
%
%  before \documentclass{sfuthesis}.
%

\usepackage{amsmath,amssymb,amsthm}
\usepackage{fixltx2e}
\usepackage[rightcaption]{sidecap}



%  FRONTMATTER  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Title page, committee page, copyright declaration, abstract,
%  dedication, acknowledgements, table of contents, etc.
%

\begin{document}

\frontmatter
\maketitle %% The details of this command are in sfuthesis.cls. Line 299
\makecommittee %% The details of this command are in sfuthesis.cls. Line 336
\makecopyrightdeclaration %% This page should stay as is, except it needs to be updated to the latest version of the declaration of partial copywrite.
				%% To update, just download the latest version, name it declaration_of_partial_copyright_licence.pdf and make sure it is saved in the
				%% Same folder as this file.

\begin{abstract}
	DNA replication starts at many sites (origins) throughout eukaryotic DNA.
	To fully understand the replication program in higher organisms, one needs to understand the behaviour of these origins.
	In \emph{Saccharomyces cerevisiae} (budding yeast), the spatial organization of the origins is simple: The origins are confined to known specific sites on the genome.
	The temporal behaviour of origins in budding yeast is more complex: They fire stochastically with a broad distribution of firing times.
	
	Several key proteins take part in the DNA replication process.
	The MCM2-7 hexamer in particular forms a helicase that unwinds DNA locally, allowing access for other proteins to replicate the separated DNA.
	Past analysis of the budding yeast replication program suggested the Multiple Initiator Model (MIM), which hypothesizes that the number of these MCMs loaded at an origin predicts the firing time of that origin.
	Part of the MIM formalism assumes that the number of loaded MCMs is large; in this case the relative fluctuations between cells will be small and are ignored.
	However, a recent experiment measuring the number of loaded MCMs has revealed that the number is low, and thus, cell-to-cell fluctuations may be larger than expected.
	The purpose of this thesis is to investigate the impact of large relative fluctuations in the number of MCMs on the MIM.
	To measure this effect, we built the ``MIM simulator,'' a modular program that simulates the replication process.
	Although a naive argument suggests that the MIM should fail when the number of MCMs is low, the impact of these fluctuations is mitigated by the contributions from neighbouring origins.
	We conclude that inferences made with the MIM remain accurate in the case that the number of MCMs is lower than first assumed.
\end{abstract}


\begin{dedication} % optional
	For anyone reading this, may you find what you're looking for.
\end{dedication}


\begin{acknowledgements} % optional
	UPDATE THIS
\end{acknowledgements}

%% making automatically managed pages: the table of contents, and lists of tables and figures.
%% To change the table of contents, check the sfuthesis.cls file. The code for the table of contents starts at line 493.
\addtoToC{Table of Contents}\tableofcontents\clearpage
\addtoToC{List of Tables}\listoftables\clearpage
\addtoToC{List of Figures}\listoffigures





%  MAIN MATTER  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Start writing your thesis --- or start \include ing chapters --- here.
%

\mainmatter%

\include{Chapters/Introduction}

\include{Chapters/Motivation}

\include{Chapters/Methods}

\include{Chapters/Results}

\include{Chapters/Conclusions}



%  BACK MATTER  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  References and appendices. Appendices come after the bibliography and
%  should be in the order that they are referred to in the text.
%

\backmatter%
	\addtoToC{Bibliography}
	\bibliographystyle{bibstyle/Thesis-Bib}%unsrt} %% default style is {plain}, but it outputs the list alphabetically by first author, unsrt will order the list by citation order
	%% Change the name ``references'' to the name of your .bib file (in the same folder as this file).
	\bibliography{references}

%% To add additional appendicies, merely add \chapters
%\begin{appendices} % optional

%\include{Appendices/EColiRep}

%\end{appendices}

% You can choose to add an index to your thesis as well. To do so, make sure you \addtoToC{Index}. You'll also need to look up how to include items in the index.
\end{document}
